---
title: "A simple analysis of Iris data"
author: "Tonje Johnsen"
format: html
editor: visual
bibliography: resources/references.bib
csl: resources/apa.csl
---

## Background

The iris dataset contains measurements of flowers from three species: setosa, versicolor, and virginica. It is available as an example dataset in R.

In this report, I explore how petal length varies between the three species.

## Research question

How does petal length vary between the three iris species?

## Data

All analyses were performed using R [@Rsoftware].

The dataset contains 150 observations, with 50 flowers from each species. Petal length is measured in centimetres.

## Results

The boxplot below shows the distribution of petal length for each of the three iris species.

```{r}
#| label: fig-petal-length
#| fig-cap: "Distribution of petal length in centimetres for three iris species. Each group contains 50 flowers."
#| echo: false

boxplot(
  Petal.Length ~ Species,
  data = iris,
  main = "Petal length by iris species",
  xlab = "Species",
  ylab = "Petal length (cm)",
  col = c("#88CCEE", "#DDCC77", "#CC6677")
)
```

Setosa has the shortest petals, while virginica generally has the longest. The distributions for versicolor and virginica overlap.

## Summary table

The table summarises petal length for each species.

```{r}
#| label: tbl-petal-summary
#| tbl-cap: "Sample size, mean and standard deviation of petal length by iris species."
#| echo: false

petal_summary <- data.frame(
  Species = levels(iris$Species),
  N = as.integer(table(iris$Species)),
  Mean = as.numeric(tapply(iris$Petal.Length, iris$Species, mean)),
  SD = as.numeric(tapply(iris$Petal.Length, iris$Species, sd))
)

knitr::kable(
  petal_summary,
  digits = 2,
  col.names = c("Species", "N", "Mean (cm)", "SD (cm)")
)
```

## References
